---
title: "rfcipDemand (Estimate Federal Crop Insurance Program Demand Models)"

output:
  github_document:
    toc: false
    toc_depth: 3
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```



```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE,
                      echo = TRUE,
                        fig.path = "man/figures/")

if (!requireNamespace("badger", quietly = TRUE)) install.packages("badger")

library(badger)

library(knitr)
library(dplyr)
library(devtools)
devtools::load_all()
devtools::document()
```


<!-- badges: start -->
[![Project Status: Active – The project has reached a stable, usable
state and is being actively
developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![Lifecycle:
experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/ftsiboe/rfcipDemand/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ftsiboe/rfcipDemand/actions/workflows/R-CMD-check.yaml)
[![codecov](https://codecov.io/gh/ftsiboe/rfcipDemand/graph/badge.svg?token=VIMHQH2SEO)](https://codecov.io/gh/ftsiboe/rfcipDemand)
![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)
![R >= 4.0](https://img.shields.io/badge/R-%3E=4.0-blue)
[![Contributor Covenant](https://img.shields.io/badge/Contributor%20Covenant-2.1-4baaaa.svg)](code_of_conduct.md)
<!-- badges: end -->

# Estimate Federal Crop Insurance Program Demand Models
 
**Acknowledgment:**  The findings and conclusions in associated publications and their associated replication files and data provided in this repository are those of the author(s) and should not be interpreted as representing the official positions, policies, or determinations of my past or current employers.

This repository centralizes research outputs, analytical tools, and resources exploring and evaluating the United States agricultural safety net programs. These government programs can be categorized into three main groups based on the type and scope of risks they address:

1. **Farm-Specific Risk Management Programs:** These programs, such as the Federal Crop Insurance Program (FCIP) and the Noninsured Crop Disaster Assistance Program (NAP), are designed to manage farm-level risks including pests and disease outbreaks, localized flooding, and equipment failure. Producers participating in these programs pay premiums for coverage.
  
2. **Regional or National Risk Coverage Programs:** These programs address widespread risks impacting multiple producers simultaneously. Notable programs include Price Loss Coverage (PLC), Agricultural Risk Coverage (ARC), and ongoing disaster assistance. Producers do not pay premiums for these types of coverage.
  
3. **Ad-Hoc Disaster Assistance Programs:** Temporary assistance programs responding to unforeseen or extraordinary events inadequately covered by existing programs, such as the COVID-19 pandemic or international trade disputes. Similar to regional programs, these do not require premium payments from producers.

**Purpose:** This toolkit aims to enhance understanding, facilitate rigorous analysis, and support informed decision-making among researchers, policymakers, students, and agricultural industry stakeholders.

**Repository Structure:** The resources in this repository are organized under thematic folders, each containing research papers/reports, analytical scripts, datasets, and reference materials.

**Intended Audience:** Agricultural economists and policy analysts; Researchers and graduate students; Policymakers and agricultural stakeholders; and Extension agents and agricultural educators

**Contributions:** Community contributions are warmly welcomed. Please feel free to open issues, submit pull requests, or provide feedback and suggestions to further enrich the resources provided here.

Go to the [**rfcipDemand wiki**](https://github.com/ftsiboe/US-FarmSafetyNet-Lab/wiki) to explore other pages in this repository

If you find it useful, please star this project and cite our papers found at [Publications co‐authored by Contributors](https://github.com/ftsiboe/US-FarmSafetyNet-Lab/wiki/Publications-co%E2%80%90authored-by-Contributors)
