---
title: "rfcipDemand (Estimate Federal Crop Insurance Program Demand Models)"

output:
  github_document:
    toc: false
    toc_depth: 3
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```



```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE,
                      echo = TRUE,
                        fig.path = "man/figures/")

if (!requireNamespace("badger", quietly = TRUE)) install.packages("badger")

library(badger)

library(knitr)
library(dplyr)
library(devtools)
devtools::load_all()
devtools::document()
```


<!-- badges: start -->
[![Project Status: Active – The project has reached a stable, usable
state and is being actively
developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![Lifecycle:
experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/ftsiboe/rfcipDemand/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ftsiboe/rfcipDemand/actions/workflows/R-CMD-check.yaml)
[![codecov](https://codecov.io/gh/ftsiboe/rfcipDemand/graph/badge.svg?token=6MKGP8Z5NB)](https://codecov.io/gh/ftsiboe/rfcipDemand)
![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)
![R >= 4.0](https://img.shields.io/badge/R-%3E=4.0-blue)
[![Contributor Covenant](https://img.shields.io/badge/Contributor%20Covenant-2.1-4baaaa.svg)](code_of_conduct.md)
[![Codecov test coverage](https://codecov.io/gh/ftsiboe/rfcipDemand/graph/badge.svg)](https://app.codecov.io/gh/ftsiboe/rfcipDemand)
<!-- badges: end -->


# Introduction
`rfcipDemand` provides tools to construct county–crop–practice–plan–unit panels from the USDA RMA Summary of Business (SOBTPU) and related sources, and to estimate FCIP demand systems with two-way cluster-robust covariance. 

The pipeline standardizes coverage measures, merges price and instrument variables, adds rental-rate and price-index controls, reconciles county acreage (FSA/NASS), and produces diagnostics including robust first-stage F-tests. Methods align with the empirical design in "The crop insurance demand response to premium subsidies Evidence from U.S. Agriculture" (Food Policy, 2023, 119(3))..

**Disclaimer:** This product uses data provided by the USDA, but is not endorsed by or affiliated with USDA or the Federal Government.

See the [LICENSE](https://github.com/ftsiboe/rfcipDemand/blob/main/LICENSE) file in the repository's root for details.

# Installation
`rfcipDemand` can be installed directly from github using 

```{r installation, echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
# Demo-only installation of rfcipCalcPass from GitHub
# (chunk is set eval=FALSE so it will not actually run)
devtools::install_github("ftsiboe/rfcipDemand",force= TRUE,upgrade = "never")
```

# Supported Functionalities


# Examples 

**NOTE:** In all examples we use .  


## Example 1: Calculations for individual-based basic crop insurance plans

This example shows how to calculate **liabilities**, **premiums**, **subsidies**, and **indemnities** using individual-based basic crop insurance plans.

We’ll use sample data for **Producer A**, who grows grain corn in **Adair County, Iowa** during the **2020 crop year**.

*Key Information*


# 📚 Citation
If you find it useful, please consider staring the repository and citing the following studies

  - Tsiboe, F. and Turner, D. (2025). [Incorporating buy‐up price loss coverage into the United States
    farm safety net](https://onlinelibrary.wiley.com/doi/full/10.1002/aepp.13536). Applied Economic Perspectives and Policy.
  - Tsiboe, F., et al. (2025). [Risk reduction impacts of crop insurance in the United States](https://onlinelibrary.wiley.com/doi/full/10.1002/aepp.13513#:~:text=In%20other%20words%2C%20on%20average,%2Dcrop%2Dyear%20revenue%20variability).
    Applied Economic Perspectives and Policy.
  - Gaku, S. and Tsiboe, F. (2024). [Evaluation of alternative farm safety net program combination strategies](https://www.emerald.com/insight/content/doi/10.1108/afr-11-2023-0150/full/html). 
  Agricultural Finance Review.

# 📬 Contact
Constructive feedback is highly appreciated, and collaborations using this package are actively encouraged. Please reach out by sending emails to Francis Tsiboe ([ftsiboe@hotmail.com](mailto:ftsiboe@hotmail.com)).

